Estimating the Effect of Urban Green Spaces on Residential House Prices in Amsterdam

### 2023-04-11 Paper/paper.Rmd
update

subparagraph: yes
header-includes: |
  \usepackage{titlesec}
  \titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \titlespacing{\subsection}{0pt}{4pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \titlespacing{\subsubsection}{0pt}{5pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \usepackage{geometry}
  \geometry{left=1.5cm,right=1.5cm}
mainfont: Cambria
bibliography: references.bib
csl: apa.csl

@@

## 2.3 Data Enrichment

To examine the economic value of urban green spaces as reflected in the real-estate prices, coordinates for each property were obtained using geocoding and transformed into a single coordinate reference system, i.e. EPSG 28992, to ensure consistency and accuracy of the analysis. Exclusion criteria were in place for properties for which coordinates could not be obtained, i.e. future housing projects. Moreover, straight line distances from each property to the nearest public amenities were enumerated using the nearest neighbour algorithm in combination with Euclidean distance. This is a common approach to adjust for the implicit premium of local public amenities on house prices [36].

@@

$$ NDVI= \frac{(N - R)}{N + R} $$

Surrounding the residential properties, mean values of NDVI were obtained at varying spatial scales, i.e. 100, 300 and 500m. In doing so, the variation of greenness around the property could be included in the analysis. Moreover, the usage of these thresholds ensured that the results were less vulnerable to scaling effects given the resolution of the data [37]. Finally, based on the extraction and enhancement of data, several independent variables were identified and constructed, as presented in Table 1. These variables were included in the hedonic pricing model, as will be elaborated on in Section 2.4, to either adjust or explain the economic value of urban green areas in Amsterdam.

@@

## 2.4 Hedonic Pricing Models

Two forms of the hedonic pricing model were employed to examine the effect of urban green spaces on housing prices per square meter, namely ordinary least squares (OLS) and geographically weighted regression (GWR). In the models, the relation of interest was adjusted for using property characteristics, transit distances and destination accessibility. Here, the usage of OLS was mainly related to its superior interpretability [38]. However, the Durbin-Watson test and Variance Inflation Factor (VIF) revealed that observations were not independent and indicated some collinearity, respectively. Hence, violating two model assumptions [39]. As a result, variables NDVI100 and NDVI500 were excluded from the analysis, as these resulted in considerable collinearity (VIF > 5). Moreover, Moran’s I indicated the presence of spatial dependency (0.590, p < 0.05), as presented in Figure 2.

@@

Furthermore, similar studies found evidence for the existence of non-stationarity in both house prices and green spaces [40, 41]. Hence, a local regression approach, i.e. GWR with an adaptive bi-square kernel, was employed to account for both the spatial dependence and non-stationarity in the data. Here, a golden search algorithm was used to obtain the kernel bandwidth, as data was not uniformly distributed across space [42]. Doing so, provided estimates for each location in the study area.

# 3. Results and Discussion

@@

# 4. Conclusion

>>>>

subparagraph: yes
header-includes: |
  \usepackage{titlesec}
  \titlespacing{\section}{0pt}{10pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \titlespacing{\subsection}{0pt}{4pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \titlespacing{\subsubsection}{0pt}{5pt plus 2pt minus 1pt}{0pt plus 1pt minus 1pt}
  \usepackage{geometry}
  \geometry{left=1.3cm,right=1.3cm}
mainfont: Cambria
bibliography: references.bib
csl: apa.csl

@@

## 2.3 Data Enrichment
  To examine the economic value of urban green spaces as reflected in the real-estate prices, coordinates for each property were obtained using geocoding and transformed into a single coordinate reference system, i.e. EPSG 28992, to ensure consistency and accuracy of the analysis. Exclusion criteria were in place for properties for which coordinates could not be obtained, i.e. future housing projects. Moreover, straight line distances from each property to the nearest public amenities were enumerated using the nearest neighbour algorithm in combination with Euclidean distance. This is a common approach to adjust for the implicit premium of local public amenities on house prices [36].

@@

$$ NDVI= \frac{(N - R)}{(N + R)} $$
  Surrounding the residential properties, mean values of NDVI were obtained at varying spatial scales, i.e. 100, 300 and 500m. In doing so, the variation of greenness around the property could be included in the analysis. Moreover, the usage of these thresholds ensured that the results were less vulnerable to scaling effects given the resolution of the data [37]. Finally, based on the extraction and enhancement of data, several independent variables were identified and constructed, as presented in Table 1. These variables were included in the hedonic pricing model, as will be elaborated on in Section 2.4, to either adjust or explain the economic value of urban green areas in Amsterdam.

@@

## 2.4 Hedonic Pricing Models
  Two forms of the hedonic pricing model were employed to examine the effect of urban green spaces on housing prices per square meter, namely ordinary least squares (OLS) and geographically weighted regression (GWR). In the models, the relation of interest was adjusted for using property characteristics, transit distances and destination accessibility. Here, the usage of OLS was mainly related to its superior interpretability [40]. However, the Durbin-Watson test and Variance Inflation Factor (VIF) revealed that observations were not independent and indicated some collinearity, respectively. Hence, violating two model assumptions [41]. As a result, variables NDVI100 and NDVI500 were excluded from the analysis, as these resulted in considerable collinearity (VIF > 5). Moreover, Moran’s I indicated the presence of spatial dependency (0.590, p < 0.05), as presented in Figure 2.

@@

  Furthermore, similar studies found evidence for the existence of non-stationarity in both house prices and green spaces [39, 42]. Hence, a local regression approach, i.e. GWR with an adaptive bi-square kernel, was employed to account for both the spatial dependence and non-stationarity in the data. Here, a golden search algorithm was used to obtain the kernel bandwidth, as data was not uniformly distributed across space [43]. Doing so, provided estimates for each location in the study area. Finally, an overview of the scripts used throughout this study may be found on: [GitHub](https://github.com/JitseRuurd/Amsterdam_house_prices.git).

# 3. Results and Discussion
  The results showed that urban green spaces provide significant economic value, indicated by their positive association with residential property prices in the local regression model. However, this association was not consistent across all locations, as presented in Figure 3. Since the local model captured both the spatial dependence and non-stationarity present in the data, this section primarily focuses on the results of this model. A detailed overview of both models is presented in Appendix I.

  Results of the local regression revealed that urban green spaces were significantly associated with residential house prices at a 300m spatial scale (p < 0.05). Here, property premiums per m2 ranged from -82,657.5 to 125,978.4 euros for each unit increase in the vegetation index. These results align with previous studies, highlighting the non-stationary nature of the association of interest [23, 36, 39, 40]. As a result, the findings of this study provide empirical evidence of the capitalisation of greenness characteristics in residential property prices.

  Moreover, the results indicated a positive association between the extent of urban green areas and residential property premiums, i.e. listings located near parks indicated higher premiums compared to smaller areas of greenness. One explanation for this finding is the direct-use value placed on these sites [23]. In other words, more extensive green areas might attract more residents to benefit from the perceived value of greenness, prompting more people to purchase houses nearby [44].

  Finally, limitations of this study were three-fold. First, distinctions between vegetation types were not considered. However, it was demonstrated to affect the scope and magnitude on house prices differently [45]. Second, the presence of property and neighbourhood characteristics that were unobserved but associated with urban green space and/or house prices may have biased the estimations. Lastly, the temporal dimension was not considered in this study. Hence, this may limit the generalisability of the study as the dynamic nature of the data might not be captured.

```{=latex}
\begin{figure}[]
  \centering
  \includegraphics[width=\linewidth]{figures/Economic_Value_Urban_Green_clip.png}
  \caption{Overview of Statistically Significant Urban Green Space Coefficients, i.e. NDVI at a 300m Spatial Scale, (p < 0.05) in Amsterdam.}
\end{figure}
```

# 4. Conclusion
  This study aimed to examine the economic value of urban green spaces in Amsterdam, as reflected in the real-estate prices of 2023. Here, property characteristics, public amenities and greenness measurements were obtained from a variety of sources. In line with topical literature, independent variables were identified and constructed to either adjust or explain the economic value of urban green spaces. In doing so, the local hedonic pricing model revealed that urban green spaces provided significant economic value. However, this association was not consistent across all locations, as the property premiums per m2 ranged from -82,657.5 to 125,978.4 euros for each unit increase in the vegetation index. The main theoretical explanation for this range might be related to the direct-use value placed on these areas. As a result, the outcomes of this study provide empirical evidence for the capitalisation of greenness characteristics in residential real-estate.

# Appendix I

@@
